---
title: "OSHA Exploratory Analysis"
output:
  html_notebook:
    highlight: pygments
    theme: spacelab
    toc: yes
  pdf_document:
    toc: yes
---
Setup - Import libraries
```{r setup, warning=FALSE, message=FALSE}
library(feather)    # for data storage
library(stringr)    # for string manipulation
library(readxl)     # for Excel import
library(lubridate)  # for date manipulation
library(scales)     # for pretty printing
library(tidyverse)  # misc
library(forcats)    # easier factor manipulation
```

The OSHA data can be found at http://ogesdw.dol.gov/views/data_summary.php.

# Convert Data

We've downloaded five data files matching with the tables in the data 
challenge and placed them in the `data` subdirectory. Now we load them in 
from CSVs using the `readr` package, which is a faster and more 
intelligent CSV importer than the base version.

Even with `readr`, this is still a slow process. CSVs are useful for 
interchange, but  aren't a great file format. A better format is 
[Apache Avro](https://avro.apache.org/), which is a standard adopted by both 
the Rstats and Python communities. We save out our data objects in this format 
using the `feather` package. We'll use only the Avro-saved objects for future 
sessions, skipping the CSV step. This is __much__ faster.

```{r convert_csvs, eval = FALSE}
load_data <- function(datatype, load_date) {
  url_base <- "http://prd-enforce-xfr-02.dol.gov/data_catalog/OSHA/osha_"
  url <- paste0(url_base, datatype, "_", load_date, ".csv.zip")
  filename <- file.path("data", paste0("osha_", datatype, ".csv.zip"))
  if (!file.exists(filename)) {
    download.file(url, filename)
  }
  unzip(filename, exdir = "data")
  read_csv(gsub(".zip", "", filename))
}

osha_load_date <- "20161019" # date of OSHA files to load in yyyymmdd format
variables <- c("inspect", "strat_codes", "optional_info", "related_activity", "violations")
sources <- c("inspection", "strategic_codes", "optional_info", "related_activity", "violation")
walk2(variables, sources, ~ {assign(.x, load_data(.y, osha_load_date), envir = .GlobalEnv)})
```

Perform some data type munging on the loaded data set, per the data dictionary and 
exploration.

```{r data_conversion, eval = FALSE}
related_activity %<>% mutate(rel_health = rel_health=="X",
                             rel_safety = rel_safety=="X")
inspect %<>% mutate(naics_code = as.integer(naics_code))
```

```{r write_feathers, eval = FALSE}
write_feather(inspect, "./data/osha_inspection.feather")
write_feather(violations, "./data/osha_violation.feather")
write_feather(strat_codes, "./data/osha_strategic_codes.feather")
write_feather(optional_info, "./data/osha_optional_info.feather")
write_feather(related_activity, "./data/osha_related_activity.feather")
```

```{r load_feathers}
inspect <- read_feather("./data/osha_inspection.feather")
strat_codes <- read_feather("./data/osha_strategic_codes.feather")
optional_info <- read_feather("./data/osha_optional_info.feather")
related_activity <- read_feather("./data/osha_related_activity.feather")
violations <- read_feather("./data/osha_violation.feather")
```

The reason for an inspection is defined in the data dictionary. The range of 
possible values is manualy copied from the data dictionary and used to set a
reference lookup value for this field.

```{r set_inspect_keys}
inspect_types <- data_frame(
  code = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M"), 
  value = c("Accident", "Complaint", "Referral", "Monitoring", "Variance", 
            "FollowUp", "Unprog Rel", "Planned", "Prog Related", 
            "Unprog Other", "Prog Other", "Other-L",  "Fat/Cat")
)
inspect_types
```

## NAICS Data

The US Census department publishes NAICS codes at http://www.census.gov/eos/www/naics/2017NAICS/2017_NAICS_Structure.xlsx. To 
get this into a referencable object, we pull down the Excel document from the 
Census site, convert it into a data frame, and perform some minimal clean up 
(conveting strings to numbers, etc.).

```{r naics}
url <- "http://www.census.gov/eos/www/naics/2017NAICS/2017_NAICS_Structure.xlsx"
filename <- "./data/2017_naics.xlsx"
if (!file.exists(filename)) download.file(url, filename, method = "libcurl", 
                                          mode = "wb")
naics <- read_excel(filename, skip = 3, 
                    col_names = c("change", "code", "naics_title", rep("blank", 3)), 
                    col_types = c("text", "text", "text", rep("blank", 3)))
naics <- naics %>% separate_rows(code, sep = "-") %>% 
  filter(is.na(code) == FALSE) %>% 
  mutate(naics_title = str_replace(naics_title, 'T$', ""), code = as.numeric(code)) %>% 
  select(-change, code, naics_title)
naics
```

# Notes on the Data

+ Data Dictionary is located at https://enforcedata.dol.gov/views/dd_display.php
+ More verbose definitions are located at https://osha.gov/oshstats/est1def.html


| Table | Object | Purpose |
|-------|--------|---------|
|Inspect | inspect | Individual parent inspection of a company |
|Violation | violations | Specific violation, (n:1) with Inspect based on `activity_nr` |
|Related Activity | relatated_activities | Links multiple inspection reports in a parent-child relationship in a activity_nr:rel_act_nr format |
|Optional Info | optional_info | Seems largely empty. Only the `opt_value` field seems to be populated with anything, and its concents seems to have no value. |
|Strategic Codes | strat_codes | TBD |

## Violation data

| Column | Notes |
|--------|-------|
|viol_type| S=Serious.W=Willful.R=Repeat.O=Other (Serious > Willful > Repeat > Other) ref (http://osha.gov/Publications/fedrites.html)|
|delete_flag| Acutal values of "X" don't match with "D" in the data dictionary. Should be removed? |


# Validate Data

Perform some sanity checks on the data.

## Inspection Data 

Valid state code include all 50  states plus D.C., VI (Virgin Islands), 
GU (Guam), PR (Puerto Rico).

```{r}
bad_state_code <- unique(inspect$site_state)[!unique(inspect$site_state) %in% 
                                               c(state.abb, "DC", "VI", "GU", "PR")]
filter(inspect, site_state %in% bad_state_code) %>% group_by(site_state) %>% tally(sort = TRUE)
```

## Violations

Violation data has a `delete _flag` which is supposed to be a "D" but 
appears to only be an "X", when present. 

```{r}
table(violations$delete_flag)
group_by(violations, delete_flag) %>% tally %>% 
  mutate(pct_of_total = percent(n / sum(n)))
```

These delete record represent a small, but non-trivial, amount of the total 
dataset.

## Strategic Codes

The startegic codes table is defined as allowing the values:
+ N=NEP (National Emphasis Program)
+ L=LEP (Local Emphasis Program)
+ S=Strategic Plan Code

```{r strat_info}
table(strat_codes$prog_type)

```

Unfortunately, there are a number of inspections that have an undefined `P` 
code. Their meaning is unclear.

```{r explore_ps}
dat <- filter(strat_codes, prog_type == "P") %>% left_join(inspect)
group_by(dat, case_mod_date) %>% tally(sort = TRUE)
```
Almost all of them are modified with a current date. Does `P` somehow indicate a pending inspection?
```{r}
dat$close_case_date %>% na.omit() %>% length()
```

Most of them have a closure date, so the theory of pending seems unfounded.

# Exploratory Analysis

With the `tibble` package, we get access to the `glimpse` function, as well
as an enhanced default print.

```{r data_overview}
glimpse(inspect)
glimpse(violations)
```

## Inspect Data

### Reasons for Insepctions
```{r inspect_reasons}
left_join(inspect, inspect_types, by = c("insp_type" = "code")) %>% 
            select(value, open_date) %>% 
  group_by(open_date = quarter(open_date, with_year = TRUE), value) %>% 
  tally %>% 
  ggplot(. , aes(x = open_date, y = n, color = fct_reorder2(value, open_date, n))) + geom_line() -> gg
gg + theme_minimal() + labs(color = "Inspection Reason")
```

Hrm. Planned audits have really plummeted. Data not current or are they not 
being performed? There's also some big spikes that look to be seasonal.

```{r inspect_seasonality}
dat <- left_join(inspect, inspect_types, by = c("insp_type" = "code")) %>% 
            select(value, open_date) %>% 
  mutate(open_date = format(open_date, "%b")) %>% 
  group_by(open_date, value) %>% 
  tally
dat$open_date <- factor(dat$open_date, levels = month.abb)
dat <- na.omit(dat)

gg <- ggplot(dat , aes(x = open_date, y = n, group = 1))
gg <- gg + stat_summary(fun.y = "sum", geom = "line", 
                        aes(group = value, color = fct_reorder2(value, open_date, n))) 
gg <- gg + scale_y_continuous(label = comma)
gg + theme_minimal() + labs(color = "Inspection Reason", 
                            title = "Monthly Totals of Inspection Reasons",
                            x = "Month",
                            y = element_blank(),
                            caption = "OSHA Inspections")
```

```{r exploratory_graphs}
dat <- inspect %>% mutate(industry = as.factor(str_match(naics_code, "(\\d{2})\\d+")[,2])) %>% 
  group_by(industry) %>% tally %>% arrange(desc(n)) %>% filter(is.na(industry) == FALSE)
gg <- ggplot(dat, aes(x = industry, y = n))
#gg <- gg + scale_y_log10()
gg <- gg + geom_bar(stat = "identity")
gg <- gg + theme_minimal()
gg
```

Industry code 23 seems super popular. What is it?

> Code 23 is: `r naics[naics$code == "23", ]$naics_title`

Ah, construction. What are all the categories in construction?

```{r}
naics %>% filter(str_detect(as.character(.$code), "^23"))
```

And how many inspections are in each of these construction categories?

```{r}
inspect %>% filter(str_detect(as.character(.$naics_code), "^23")) %>% 
  group_by(naics_code, inspect_year = year(open_date)) %>% summarize(inspections = n()) %>% 
  mutate(pct_by_year = percent(inspections / sum(inspections))) %>% 
  arrange(desc(inspect_year), desc(inspections)) %>% 
  left_join(naics, by = c("naics_code" = "code")) %>% 
  select(naics_title, inspect_year, naics_code, everything())

```

So commerical and roofing are highly inspected industries.
```{r}
dat <- inspect %>% mutate(industry = str_match(naics_code, "(\\d{2})\\d+")[,2]) %>% 
  group_by(industry) %>% tally %>% arrange(desc(n)) %>% filter(is.na(industry) == FALSE)
dat <- left_join(dat, mutate(naics, code = as.character(code)), by = c("industry" = "code")) %>% 
  group_by(naics_title) %>% summarize(n = sum(n))
gg <- ggplot(dat, aes(x = fct_reorder(factor(naics_title), n), y = n))
#gg <- gg + scale_y_log10()
gg <- gg + geom_bar(stat = "identity")
gg <- gg + labs(x = "Industry", y = "Number of Inspections", title = "Inspections per Industry")
gg <- gg + scale_y_continuous(label = comma) + coord_flip()
gg <- gg + theme_minimal()
gg
```

### Advanced Notices

How have advanced notices of inspections changed over time?

```{r adv_notices}
dat <- inspect %>% mutate(yr = year(open_date)) %>% 
  group_by(yr, adv_notice) %>% tally() %>% na.omit()
gg <- ggplot(dat, aes(x = yr, y = n, color = adv_notice)) + geom_line() + 
  theme_minimal() + labs(title = "Change in Advanced Notice of Inspections over time", 
                         caption = "Source: OSHA dataset",
                         x = "Year",
                         y = "Number of Inspections") +
  scale_y_continuous(labels = comma)
gg
```

Interesting. It looks like advanced notice wasn't tracked, or not given, prior 
to about 1982. There's also a big drop of total inspections at the end. We 
should look at the distribution of inspections over time.

```{r total_inspections}
dat <- inspect %>% 
  group_by(open_date = quarter(open_date, with_year = TRUE)) %>% 
  tally

#gg <- ggplot(inspect, aes(x=open_date, y = n, group=quarter(open_date, with_year = TRUE))) + geom_line(na.rm = TRUE)
#gg <- gg + scale_x_date(date_breaks = "6 months")

gg <- ggplot(dat, aes(x = open_date, y = n)) + geom_line(na.rm = TRUE)
gg <- gg + labs(title = "Inspections over Time",
                subtitle = "Grouped by Quarter",
                caption = "OSHA Inspection data",
                x = "Date",
                y = "Number of Inspections")
gg + theme_minimal() + geom_smooth(method = "loess", na.rm = TRUE)
```

Yup. Inspections are down from the late 80s. They've been more or less 
holding steady over the past decade, though there could be a decreasing 
trend. The big drop of inspections from 77 - 82 is interesting.

## Violations

Violations have an initial penalty (assessed at the time of report authoring) 
and a current penalty (presumably a negotiated settlement). Let's look at these

```{r violation_amounts}
dat <- violations %>%  select(initial_penalty, current_penalty) %>% 
  na.omit() %>% filter(initial_penalty > 5473 ) %>% 
  mutate(penalty_change = current_penalty - initial_penalty) %>% 
  gather(penalty_type)

dat$penalty_type <- factor(dat$penalty_type, levels=c("initial_penalty", 
                                                      "current_penalty", "penalty_change"))
gg <- ggplot(filter(dat, as.character(penalty_type) != "penalty_change"), aes(x = value)) + geom_histogram(bins = 50)
#geom_histogram()
#+ facet_grid(penalty_type ~ .)
gg <- gg + scale_x_continuous(trans = "log1p", label=dollar) + facet_grid(penalty_type ~ .)
gg <- gg + labs(title = "Initial and Current Penalties", caption = "OSHA Violations")
gg + theme_minimal()

gg <- ggplot(filter(dat, as.character(penalty_type) == "penalty_change"), aes(x = penalty_type, y = value)) 
gg <- gg + geom_boxplot()
gg <- gg + scale_y_continuous(label = dollar)
gg <- gg + labs(title = "Change in Current vs. Initial Penalties", caption = "OSHA Violations")
gg + theme_minimal()

```
