---
title: "Redshift test"
output: html_notebook
---

```{r setup}
library(dplyr)
library(RPostgreSQL)  # no TLS support

# dplyr only works with RPostgreSQL at the moment
# devtools::install_github("RcppCore/Rcpp")
# devtools::install_github("rstats-db/DBI")
# devtools::install_github("rstats-db/RPostgres")
#library(RPostgres)
```


```{r connect_to_redshift}
pg_dsn <- paste0(
    'dbname=', 'osha', ' ',
    'sslmode=require'
)
host <-  'tf-redshift-cluster.cx8afkkusjuc.us-east-1.redshift.amazonaws.com'
port <- 5439
username <- 'mcp'
password <- Sys.getenv('RedShift_password')
con <- dbConnect(RPostgres::Postgres(),
          dbname = 'osha', 
          host = host, 
          port = port, 
          password = password, 
          user = username)

```

# Load CSVs into Redshift

Get ARN of role to use for reading from S3.

```{r get_s3_reader_arn}
cmd <- paste0("aws sts get-caller-identity --output text --query Account")
aws_account_id <- system(cmd, intern = TRUE)

role_name <- c("redshift_s3_reader")

credentials <- paste0("aws_iam_role=arn:aws:iam::", aws_account_id, ":role/", role_name)
message(paste0("Credentials for S3 Reader are: ", credentials))
```

Create a helper function to determine the appropriate SQL structure for a 
data_frame.

```{r determine_table_structure}
find_str <- function(dat) { dat %>% map_df(., 
                                           ~ data_frame(length = max(nchar(.x), na.rm=TRUE), 
                                                        classtype = class(.x)), 
                                           .id = "column_name")
}
tables <- c("optional_info", "related_activity", "strat_codes", "violations")
```

## Inspect
```{sql create_violations_table, connection = con}
CREATE table violation(
  activity_nr       integer not null,
  citation_id       varchar(7),
  delete_flag       varchar(1),
  standard          varchar(33),
  viol_type         varchar(1),
  issuance_date     date,
  abate_date        date,
  abate_complete    varchar(1),
  current_penalty   decimal,
  initial_penalty   decimal,
  contest_date      date,
  final_order_date  date,
  nr_instances      integer,
  nr_exposed        integer,
  rec               varchar(1),
  gravity           varchar(2),
  emphasis          varchar(1),
  hazcat            varchar(10),
  fta_insp_nr       varchar(9),
  fta_issuance_date varchar(10),
  fta_penalty       varchar(10),
  fta_contest_date  varchar(10),
  fta_final_order_date varchar(10),
  hazsub1           varchar(4),
  hazsub2           varchar(4),
  hazsub3           varchar(4),
  hazsub4           varchar(4),
  hazsub5           varchar(4),
  load_dt           datetime,
  PRIMARY KEY(activity_nr)
  )
distkey(activity_nr)
```
```{sql create_strategic_codes_table, connection = con}
CREATE table strategic_codes(
  activity_nr       integer not null,
  prog_type         varchar(1),
  prog_value        varchar(25),
  load_dt           datetime,
  PRIMARY KEY(activity_nr)
  )
distkey(activity_nr)
```

```{sql create_related_activity_table, connection = con}
CREATE table related_activity(
  activity_nr       integer not null,
  rel_type          varchar(1),
  rel_act_nr        integer,
  rel_safety        varchar(1),
  rel_health        varchar(1),
  load_dt           datetime,
  PRIMARY KEY(activity_nr)
  )
distkey(activity_nr)
```

```{sql create_optional_info_table, connection = con}
CREATE table optional_info(
  activity_nr       integer not null,
  opt_type          varchar(1),
  opt_id            integer,
  opt_value         varchar(50),
  opt_info_id       varchar(1),
  load_dt           datetime,
  PRIMARY KEY(activity_nr)
  )
distkey(activity_nr)
```

```{sql create_inspect_table, connection = con}
CREATE table inspect(
  activity_nr       integer not null,
  reporting_id      integer,
  state_flag        varchar,
  estab_name        varchar(100),
  site_address      varchar(142),
  site_city         varchar(30),
  site_state        varchar(18),
  site_zip          varchar(8),
  owner_type        varchar(5),
  owner_code        int,
  adv_notice        varchar(4),
  safety_hlth       varchar(8),
  sic_code          int,
  naics_code        int,
  insp_type         varchar(6),
  insp_score        varchar(6),
  why_no_insp       varchar(1),
  union_status      varchar(1),
  safety_manuf      varchar(1),
  safety_const      varchar(1),
  safety_marit      varchar(1),
  health_manuf      varchar(1),
  health_const      varchar(1),
  health_marit      varchar(1),
  migrant           varchar(1),
  mail_street       varchar(110),
  mail_city         varchar(30),
  mail_state        varchar(2),
  mail_zip          varchar(5),
  host_est_key      varchar(18),
  nr_in_estab       int,
  open_date         date,
  case_mod_date     date,
  close_conf_date   date,
  close_case_date   date,
  ld_dt             timestamp,
  PRIMARY KEY(activity_nr)
  )
distkey(activity_nr)
```

```{sql, connection = con, output.var = sql_result}
COPY INSPECT
  FROM 's3://users-severski/davidski/osha/osha_inspection.csv.bz2' 
  REGION 'us-west-2'
  credentials ?credentials
  CSV IGNOREHEADER AS 1
  BZIP2
  MAXERROR 1000;
```

Load the `violation` table from CSV.

```{sql, connection = con, output.var = sql_result}
COPY VIOLATION
  FROM 's3://users-severski/davidski/osha/osha_violation.csv.bz2' 
  REGION 'us-west-2'
  credentials ?credentials
  CSV IGNOREHEADER AS 1
  BZIP2
  MAXERROR 1000;
```

Load the `related_activity` table from CSV.

```{sql, connection = con, output.var = sql_result}
COPY RELATED_ACTIVITY
  FROM 's3://users-severski/davidski/osha/osha_related_activity.csv.bz2' 
  REGION 'us-west-2'
  credentials ?credentials
  CSV IGNOREHEADER AS 1
  BZIP2
  MAXERROR 1000;
```

Load the `strategic_codes` table from CSV.

```{sql, connection = con, output.var = sql_result}
COPY STRATEGIC_CODES
  FROM 's3://users-severski/davidski/osha/osha_strategic_codes.csv.bz2' 
  REGION 'us-west-2'
  credentials ?credentials
  CSV IGNOREHEADER AS 1
  BZIP2
  MAXERROR 1000;
```

Load the `optional_info` table from CSV.

```{sql, connection = con, output.var = sql_result}
COPY OPTIONAL_INFO
  FROM 's3://users-severski/davidski/osha/osha_optional_info.csv.bz2' 
  REGION 'us-west-2'
  credentials ?credentials
  CSV IGNOREHEADER AS 1
  BZIP2
  MAXERROR 1000;
```

Check the created table structure

```{sql, connection = con, output.var = sql_result}
SELECT "column", type, encoding, distkey, sortkey
from pg_table_def where tablename = '';
```

## Check Results

Show our results.

```{r}
sql_result
```
Show the most recent errors

```{sql check_load, connection = con}
select starttime, filename, err_reason, line_number,
  colname, type, col_length, position, raw_field_value,
  raw_line, err_code
from stl_load_errors
order by starttime desc;
```

Show the first 100 records from the RedShift table

```{sql connection = con}
select TOP 100 * from inspect;
```

Doing the same thing, but with R code....
```{r dbi_query}
dat <- dbGetQuery(con, "select TOP 100 * from inspect;")
dat
```

```{sql connection = con, exec=FALSE}
DROP TABLE violations;
```

# Connect to data with dplyr

```{r dplyr_setup}
src <- src_postgres(
  dbname = 'osha',
  host = host,
  port = port,
  password = password,
  user = username)
```

Get list of tables. Connect to the inspect table.

```{r dplyr}
src_tbls(src)       # list all tables
dat <- tbl(src, "inspect") # connect to the inspect table
dat
```

Sample query with dplyr

```{r query_with_dplyr}
filter(dat, site_state == "WA") %>% 
  mutate(inspect_year = DATE_PART('year', open_date)) %>% 
  group_by(inspect_year, site_city) %>% 
  tally %>% 
  filter(cume_dist(n) > 0.5) %>% 
  #mutate(freq = n /sum(n)) %>%  
  arrange(desc(inspect_year), desc(n)) %>% collect
```


